---
title: "LLM_Logic"
author: "Carson Chiem"
date: "2024-01-22"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# importing packages and data
library(tidyverse)
library(ordinal)
dataset = read_csv("C:/CS_programs/Python/LLM_Logic/data/output_responses.csv")
```

```{r}
# preparing dataframe
# filtering out control trials
df <- dataset %>%
  filter(Condition != "Control")

#dataset$Continuation[dataset$Condition == "Control"] = "Control"
```

```{r}
# perform factorial ANOVA
anova_result <- anova(aov(Response ~ Continuation, data = df))

print(anova_result)
```
Type: The p-value associated with the "Type" variable is slightly greater than the significance level, indicating that there is no statistically significant difference in the mean responses among the different types. 

Continuation: The p-value associated with the "Continuation" variable is less than the significance level, indicating that there is a statistically significant difference in the mean responses between different continuation conditions.

Type:Continuation: The p-value associated with the interaction between "Type" and "Continuation" is greater than the significance level, indicating that there is no statistically significant interaction effect between these two variables.

Overall, based on the p-values, we can conclude that "Continuation" has significant effects on the response variable. "Type" has no significant effects on the response, and there is no significant interaction between them as well.

```{r}
# using Cumulative Linked Mixed Model (used in original study)
# specifying variables as factors
df$Response = factor(df$Response)
df$Type = factor(df$Type)

# fit Cumulative Link Mixed Model
clmm_model <- clmm(Response ~ Type * Continuation + (1|Type) + (1|Continuation), data = df)

summary(clmm_model)
```
The NOR continuation, corresponding to the narrow scope in the original study, is a statistically significant predictor of the response variable with a p-value of 0.0127. This is not the case in the original study, where they found the "NAND"/wide scope to be the predictor. 

The Type "Other DE" is also a significant predictor of the response variable.
