---
title: "LLM_exp1"
author: "Carson Chiem"
date: "2025-01-13"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
# Importing packages and data
library(tidyverse)
library(ordinal)
dataset = read_csv("output_responses_llama_3.2_3b_prompting_exp1_1.csv")
View(dataset)
```
```{r}
# Convert relevant columns to factors for clarity
dataset$Contradiction <- as.factor(dataset$Contradiction)
dataset$Response <- as.factor(dataset$Response)
dataset$Connective <- as.factor(dataset$Connective)
```

Instructions:  
1: "Is the following sentence a contradiction?", Contradiction 1 and output yes is good  
2: "Does the following sentence make sense?", Contradiction 1 and output no is good  
3: "Is this statement logically valid?", Contradiction 1 and output no is good  

```{r}
dataset <- dataset %>%
  # Standardize the Response column
  mutate(Response = case_when(
    str_detect(Response, regex("\\bYes\\b", ignore_case = TRUE)) ~ "Yes",
    str_detect(Response, regex("\\bNo\\b", ignore_case = TRUE)) ~ "No",
    TRUE ~ Response  # Leave other responses as they are (optional, for debugging purposes)
  )) %>%
  # Checking if response is correct
  mutate(Correct = case_when(
    # Logic for Instructions == 1
    Instructions == 1 & Contradiction == 1 & Response == "Yes" ~ 1,
    Instructions == 1 & Contradiction == 0 & Response == "No" ~ 1,
    
    # Opposite logic for Instructions == 2 or 3
    (Instructions == 2 | Instructions == 3) & Contradiction == 1 & Response == "No" ~ 1,
    (Instructions == 2 | Instructions == 3) & Contradiction == 0 & Response == "Yes" ~ 1,
    
    # Incorrect otherwise
    TRUE ~ 0
  ))
```

```{r}
experimental_data <- dataset %>% filter(Type == "experimental")
control_data <- dataset %>% filter(Type == "control")
```

Connectives:  
- and, if, so, but, although, when, and therefore create contradictions  
- or, maybe, and perhaps do not create contradictions  
```{r}
# Calculating accuracy for experimental data
# Calculate accuracy for each connective
accuracy_exp_by_connective <- experimental_data %>%
  group_by(Connective) %>%
  summarise(
    Total = n(), 
    Correct = sum(Correct), 
    Accuracy = round((Correct / Total) * 100, 2)
  )

print(accuracy_exp_by_connective)

# Calculate accuracy for each instruction 
accuracy_exp_by_instructions <- experimental_data %>%
  group_by(Instructions) %>%
  summarise(
    Total = n(), 
    Correct = sum(Correct), 
    Accuracy = round((Correct / Total) * 100, 2)
  )

print(accuracy_exp_by_instructions)

# Calculate accuracy by both connectives and instructions
accuracy_exp_by_conn_inst <- experimental_data %>%
  group_by(Connective, Instructions) %>%
  summarise(
    Total = n(),
    Correct = sum(Correct),
    Accuracy = round((Correct / Total) * 100, 2)
  )

print(accuracy_exp_by_conn_inst)
```
Observations 3.2 1b:  
- In table 1, the connectives that do not create contradictions (or, maybe, might, perhaps)have the lowest accuracies on average (with the exception of "if")  
- In table 2, the 2nd instruction "Does the following sentence make sense?" yielded the most accurate responses  
- In table 3, the connectives that do not create contradictions (or, maybe, might, perhaps) respond more accurately when asked the 1st instruction "Is the following sentence a contradiction?"  
- In table 3, the connectives that create contradictions respond more accurately when asked the 2nd instruction "Does the following sentence make sense?"  

Observations 3.2 3b:  
- In table 1,the connectives that do not create contradictions (or, maybe, might, perhaps)have the lowest accuracies on average, while the other are over 90%  
- In table 2, the 2nd instruction "Does the following sentence make sense?" yielded the most accurate responses  
- In table 3, the connectives "maybe", "perhaps", and "might" performed the best with instruction 2, and terribly with instruction 1
- In table 3, with the connectives that create contradictions, there is no one instruction that they perform better or worse than in general (very random)  

```{r}
# Calculating accuracy for control data
# Calculate accuracy for each connective
accuracy_con_by_connective <- control_data %>%
  group_by(Connective) %>%
  summarise(
    Total = n(), 
    Correct = sum(Correct), 
    Accuracy = round((Correct / Total) * 100, 2)
  )

print(accuracy_con_by_connective)

# Calculate accuracy for each instruction
accuracy_con_by_instructions <- control_data %>%
  group_by(Instructions) %>%
  summarise(
    Total = n(), 
    Correct = sum(Correct), 
    Accuracy = round((Correct / Total) * 100, 2)
  )

print(accuracy_con_by_instructions)

# Calculate accuracy by both connectives and instructions
accuracy_con_by_conn_inst <- control_data %>%
  group_by(Connective, Instructions) %>%
  summarise(
    Total = n(),
    Correct = sum(Correct),
    Accuracy = round((Correct / Total) * 100, 2)
  )

print(accuracy_con_by_conn_inst)
```

Observations 3.2 1b:  
- In table 1, all connectives have low accuracy scores  
- In table 2, accuracy rates are much higher for 1st instruction compared to 2nd or 3rd instructions( possibly because our control sentences don't make much sense in the conventional sense despite being non-contradictory sentences)  
- Table 3 supports the observations from table 2  

Observations 3.2 3b:  
- In table 1, all connectives have low to middle accuracy scores  
- In table 2, accuracy rates are much higher for 2nd instruction compared to 1st or 3rd instructions  

Statistics:  
P < 0.05 means the model’s performance likely differs from chance  

```{r}
# Contingency table for each connective
p_values_connective <- experimental_data %>%
  group_by(Connective) %>%
  summarize(
    Total_Correct = sum(Correct),
    Total_Incorrect = n() - sum(Correct)
  ) %>%
  rowwise() %>%
  mutate(
    P_Value = chisq.test(matrix(c(Total_Correct, Total_Incorrect), nrow = 2))$p.value
  )

print(p_values_connective)

# Contingency table for each instructions type
p_values_instructions <- experimental_data %>%
  group_by(Instructions) %>%
  summarize(
    Total_Correct = sum(Correct),
    Total_Incorrect = n() - sum(Correct)
  ) %>%
  rowwise() %>%
  mutate(
    P_Value = chisq.test(matrix(c(Total_Correct, Total_Incorrect), nrow = 2))$p.value
  )

print(p_values_instructions)
```
Observations 3.2 1b:  
- all of the connectives that create contradictory statements are statistically significant except "although", "if", and "when"  
- The 3rd instruction, "Is this statement logically valid?" is not statistically significant  

Observations 3.2 3b:  
- all of the connectives that create contradictory statements are statistically significant, while those that don't are not statistically significant
- All instructions are statistically significant
```{r}
# Fit logistic regression model with interaction term
interaction_model <- glm(
  Correct ~ Connective * Instructions, 
  data = dataset, 
  family = binomial(link = "logit")
)

# Summary of the model
summary(interaction_model)
```
Observations 3.2 1b:  
- the only statistically significant connective is "maybe" when interacted with instructions  

Observations 3.2 3b:  
- the only statistically significant connectives are "and", "but", and "so" when interacted with instructions  

```{r}
# Bar plot for accuracy by connective
library(ggplot2)

ggplot(accuracy_exp_by_connective, aes(x = Connective, y = Accuracy, fill = Connective)) +
  geom_bar(stat = "identity") +
  labs(
    title = "Accuracy by Connective (Experimental Data)",
    x = "Connective",
    y = "Accuracy (%)"
  ) +
  theme_minimal() +
  theme(legend.position = "none")

```

```{r}
# Bar plot for accuracy by instructions
ggplot(accuracy_exp_by_instructions, aes(x = factor(Instructions), y = Accuracy, fill = factor(Instructions))) +
  geom_bar(stat = "identity") +
  labs(
    title = "Accuracy by Instructions (Experimental Data)",
    x = "Instruction Type",
    y = "Accuracy (%)"
  ) +
  theme_minimal() +
  theme(legend.position = "none")
```

```{r}
# Calculate mean accuracy and standard error for each connective
accuracy_with_se <- experimental_data %>%
  group_by(Connective) %>%
  summarise(
    MeanAccuracy = mean(Correct) * 100,  # Convert to percentage
    SE = sd(Correct) / sqrt(n()) * 100  # Standard error as percentage
  )

# Plot with error bars
ggplot(accuracy_with_se, aes(x = Connective, y = MeanAccuracy, fill = Connective)) +
  geom_bar(stat = "identity", color = "black") +
  geom_errorbar(aes(ymin = MeanAccuracy - SE, ymax = MeanAccuracy + SE), width = 0.2) +
  theme_minimal() +
  labs(
    title = "Mean Accuracy by Connective with Error Bars",
    x = "Connective",
    y = "Mean Accuracy (%)"
  )
```

